### 2022-03-04 paper1.tex
finish paper 1

\section{Abstract}
I will first investigate how Google Search Trends can be operationalized
as cultural attitudes, disease prevalence and voting behaviors through
comparing trends data to observed valid indicators of these cases.
Specifically, I will test cultural indicators from the NORC General
Social Survey, United States county-level suicide rates from National
Center for Health Statistics, US case rates of Covid-19 from The New
York Times, historical US Presidential Election results, and the
American National Election Survey against selected Google Trends terms
and analyze the data using correlation methods. The aim of this article
is to assess the construct validity and will contribute to critical big
data studies and sociological methodology.

@@

One source of big data that has yet to be widely utilized are Google Search Trends. 
Google Trends is an open data resource provided by Google, a major internet search
engine worldwide, that analyzes the top search queries across different localities, languages, and time. 
While Google Trends is freely available and holds a breadth of online activities that could potentially 
be of use in social science, few social scientists have used this resource. Because of this, 
I set out to establish how can we operationalize Google Search Trends as
validated indicator for uses in social science research to provide guidance to social scientists as they
adapt different sources of big data into their toolboxes.

@@

Google Search Trends have been used continuously in estimations of
disease prevalence and population health in journals like the Journal of
Medical Internet Research. While much of this research has focused on
the Covid-19 pandemic \citep{jimenez_etal20, jimenezCOVID19SymptomGoogle2020, limEstimatingInformationSeekingBehaviour2020, mavraganiCOVID19PredictabilityUnited2020, nguyenGoogleTrendsAnalysis2020, todorovaInternetBasedData2021},
other research has investigated Google Trends as indicators of wellbeing
\citep{brodeurCOVID19LockdownsWellbeing2021, carpiTwitterSubjectiveWellBeing2020, duCOVID19IncreasesOnline2020},
suicidality \citep{burnettTimeTrendsPublic2020}, 
vaccination uptake \citep{dalumhansenEnsembleLearnedVaccination2016}, 
obesity \citep{sarigulNowcastingObesityUsing2014}, 
and even insomnia \citep{zittingGoogleTrendsReveal2020},
to cover a few examples. For a partial review of other utilizations, see \citet{nutiUseGoogleTrends2014}. According to
\citet{jaidkaInformationseekingVsSharing2021}, the majority of studies profess
a correlation of \> .70, "demonstrating the vast potential of Google
Search as a proxy for monitoring population health" (p. 3) based on
assumptions that individuals search because of self-diagnosis and to
identify possible courses of treatment \citep{dechoudhurySeekingSharingHealth2014}.

\citet{mingUnderstandingHealthCommunication2021}

@@

\subsection{Causes for concern}
summarize
\citet{jungherr_etal17}

What constructs might google trends capture and not capture well? Capture
attention but not attitudes

In contrast to the papers mentioned above, there are researchers who have cautioned against
the wide usage of Google Search Trends. For instance, \citet{asseoTrackingCOVID19Using2020} 
concludes in their research on Covid-19 case rates and Google searches that the correlation 
is often spurious and due to confounding. They even explain that "searches would
result not only from self-symptoms, but also from interest elicited by
media coverage" \citep[][TODO ADD Page here]{asseoTrackingCOVID19Using2020}.

\citet{rovetta21}

\citet{cebrian_domenech22}

@@

\subsubsection{Google Trends}
This paper focuses on a validation of Google search trends \citep{googletrends}.
Google trends portray the search frequency for specific search terms across
designated media markets areas (DMAs), a nonoverlapping aggregation of U.S.
counties to 210 media markets based on similar population clusters \citep{dma_key}.
Raw data is on a scale from 0 to 100, with 100 being the maximum search
popularity out of all DMAs. When available, I use Google search topics to
measure trends as opposed to search terms. Search topics are a more robust
measurement than a single search term: topics are aggregations of the rates of
multiple, highly correlated search terms together into a cohesive topic. For
example, while 'Beyoncé', 'Beyonce' and 'beyonce knowles' are all separate
search terms, 'Beyoncé Knowles' encompasses all of these into a single search
topic. Google Search Trend are only available cross-sectionally (a single time
period across a geography) or as time-series (a single geo-location across
time). To remedy this and build a longitudinal dataset of each search topic for
the longitudinal datasets, I follow the method proposed in \citet[p. 5]{park_etal}.
This method involves building a dataset of unscaled cross-sectional values,
selecting a DMA to use to establish the rescaling ratio (I use 'Los Angeles
CA'), and then finding the time-series values for the one DMA. To find the
rescaling ratio for each week in the time-series, you divide the time-series
value for each week by the cross-sectional value for each week, resulting in a
rescaling vector to be used for all weeks in the dataset across geographies. To
rescale each longitudinal value, multiply the respective week's rescaling ratio
by the cross-sectional value. Rescaled longitudinal data was compared against
time-series data for multiple test counties and was equivalent. For a more
in-depth explanation of this procedure, see \citet[p. 5]{park_etal}. Missing
datapoints in longitudinal datasets were filled in with interpolated values
using `zoo::na.approx()` \citep{zoo}.

@@

\subsection{Analysis}
Google Search Trends data and additional demographic data are merged with each
individual indicator based on County FIPS Codes and date. After creating these
different datasets, I use the Pearson correlation formula (formula \eqref{eq:pearsoncorr}
) for cross-sectional numeric data to calculate the
strength of the relationship between each Google Trend and the respective
data source.

@@

To address longitudinal correlations, I employ Repeated Measures Correlation
using the `rmcorr` package in R \citep{bland1995, bakdash2017}. Repeated Measures
Correlation is useful for determining the within-county association for paired
measures across time and counties. Results of these correlations can be seen in 
table \ref{tab:corr-results}.

As an additional test of the relationships, I employ multiple linear regression
(for cross-sectional datasets) and random intercept hierarchical linear models
\citep{pinheiro_etal21} (for longitudinal datasets) to identify the strength of
relationships across locales: I include city-data like county population size,
broadband rates, and median income to attempt to disentangle possible
confounders of the relationship between Google search Trends and verified
indicators. Identifying these possible confounded relationships will help to
explain why some articles find relationships between the trends and outcomes
while others did not. For linear regression models, I normalize independent
variables, i.e.variables have been centered and scaled to have a mean of 0 and
standard deviation 1.

>>>>

\section{Abstract}

Computational social science and other disciplines are quick to 
adopt new sources of digital trace data for use in academic research. 
In this article, I examine the validity of Google Search Trends 
as potential indicator of three different cases, namely attitudes, 
disease prevalence and political preferences using eight %TODO double check number
different validated data sources. I use Pearson and Repeated Measures Correlations 
between the Google Trends and the validated indicators as well as multiple linear regression
(for cross-sectional datasets) and random intercept hierarchical linear models
(for longitudinal datasets) as an additional test of the data. 
I fail to find moderate or strong correlation among any of the Google 
Trends tested and their validated indicators. While some Google Trends 
tested were significantly associated with the outcome in the regression models,
effects tended to be small and the total model explainability remained low,
even when controlling for demographic variables. This article shows that there is no 
no external validity of Google Trends for these uses and social scientists
will find no replacement for high quality survey data with Google Trends.

\textbf{Keywords}: Social Science Methodology, Digital Trace Data, Information-Seeking Behavior, Google Trends, Information and communications technology, Validity

@@

One source of big data that has yet to be widely utilized are Google Search Trends. 
Google Trends is an open data resource provided by Google, a major internet search
engine worldwide, that analyzes the top search queries across different localities, 
languages, and time. While Google Trends is freely available and holds a breadth of 
online activities that could potentially be of use in social science, few social 
scientists outside of economics \citep[see][for examples]{choi2012predicting, jun2018ten,da2011search} have used this resource.
Because of this, I set out to establish how can we operationalize Google Search Trends as
validated indicator for uses in social science research to provide guidance to social 
scientists as they adapt different sources of big data into their toolboxes.

@@

Google Search Trends have been used continuously in estimations of
disease prevalence and population health in journals like the Journal of
Medical Internet Research. While much of this research has focused on
the Covid-19 pandemic \citep{jimenez_etal20, jimenezCOVID19SymptomGoogle2020, limEstimatingInformationSeekingBehaviour2020, mavraganiCOVID19PredictabilityUnited2020, nguyenGoogleTrendsAnalysis2020, todorovaInternetBasedData2021, mingUnderstandingHealthCommunication2021},
other research has investigated Google Trends as indicators of wellbeing
\citep{brodeurCOVID19LockdownsWellbeing2021, carpiTwitterSubjectiveWellBeing2020, duCOVID19IncreasesOnline2020},
suicidality \citep{burnettTimeTrendsPublic2020}, 
vaccination uptake \citep{dalumhansenEnsembleLearnedVaccination2016}, 
obesity \citep{sarigulNowcastingObesityUsing2014}, 
and even insomnia \citep{zittingGoogleTrendsReveal2020},
to cover a few examples. For a partial review of other utilizations, see \citet{nutiUseGoogleTrends2014}. According to
\citet{jaidkaInformationseekingVsSharing2021}, the majority of studies profess
a correlation of \> .70, "demonstrating the vast potential of Google
Search as a proxy for monitoring population health" (p. 3) based on
assumptions that individuals search because of self-diagnosis and to
identify possible courses of treatment \citep{dechoudhurySeekingSharingHealth2014}.

@@

\subsection{Causes for concern}

While Google Trends has not been widely used or tested in social science,
I am not the first author to investigate its validity.
In contrast to the papers mentioned above, there are researchers who have cautioned against
the wide usage of Google Search Trends. For instance, \citet{asseoTrackingCOVID19Using2020} 
concludes in their research on Covid-19 case rates and Google searches that the correlation 
is often spurious and due to confounding. They even explain that "searches would
result not only from self-symptoms, but also from interest elicited by
media coverage" \citep[][p.1]{asseoTrackingCOVID19Using2020}.

\citet{rovetta21} also investigate the reliability of Covid-19 searches to investigate the quality of the data and find that there are various issues with missing data in smaller regions and significant variation in correlations between regions throughout the study period, which they indicated was a cause for concern more than a product of behavioral changes.

\citet{cebrian_domenech22} emphasize in their research that there are some 
non-negligible issues with the measurement of Google Trends, pointing to 
errors within the purview of accuracy, completeness, consistency, and validity \citep{KarrDataQuality}. \citet{cebrian_domenech22} point not only to 
meta issues like the lack of transparency into the sampling methods 
and population coverage, but they identify real differences 
in the data provided by the Google Trends API for the same query, pointing
to likely internal process used by Google to sample their data to compute the trends.
These authors show that the variability of the results over time for the same query
can alter the results of analysis and forecasting significantly; they propose
caution in the use of the digital trace data for other researchers until
further research can determine the cause for the inaccuracies.

Finally, it is important to note that a majority of research using 
Google Trends as digital trace data do so to indicate
populations \textit{attention} to a search query. Because
searches can be done with many different intentions, such as 
information gathering, ridicule, navigational purposes, it may
be hard to declare that an overall level of engagement with a topic
on Google Search may be indicative of anything more than interest \citep{da2011search}. 
\citet{jungherr_etal17} explore this topic deeply with data from
Twitter on German federal election candidates and find that there is no
indication that tweets about candidates are indicative of or can predict 
political support. Instead, they conclude that tweets merely indicate the
temporal dynamics of public attention toward politics.

Building on this previous research in the reliability and accuracy of the trends, 
it is imperative that, as a discipline, we are able to provide a framework for
how to use different sources of digital trace data. This aim of this paper is
to take the first step to build that framework. I will attempt to validate 
Google Search Trends for use outside of mere interest or attention using the
methods that follow; specifically, to verify whether
trends can be used to indicate attitudes, health outcomes, or political support.

@@

\subsubsection{Google Trends}
This paper focuses on a validation of Google search trends \citep{googletrends}.
Google trends portray the search frequency for specific search terms across
designated media markets areas (DMAs), a nonoverlapping aggregation of U.S.
counties to 210 media markets based on similar population clusters \citep{dma_key}.
Raw data is on a scale from 0 to 100, with 100 being the maximum search
popularity out of all DMAs. When available, I use Google search topics to
measure trends as opposed to search terms. Search topics are a more robust
measurement than a single search term: topics are aggregations of the rates of
multiple, highly correlated search terms together into a cohesive topic. For
example, while 'Beyoncé', 'Beyonce' and 'beyonce knowles' are all separate
search terms, 'Beyoncé Knowles' encompasses all of these into a single search
topic. Google Search Trend are only available cross-sectionally (a single time
period across a geography) or as time-series (a single geo-location across
time). To remedy this and build a longitudinal dataset of each search topic for
the longitudinal datasets, I follow the method proposed in \citet[p. 5]{park_etal}.
This method involves building a dataset of unscaled cross-sectional values,
selecting a DMA to use to establish the rescaling ratio (I use 'Los Angeles
CA'), and then finding the time-series values for the one DMA. To find the
rescaling ratio for each week in the time-series, you divide the time-series
value for each week by the cross-sectional value for each week, resulting in a
rescaling vector to be used for all weeks in the dataset across geographies. To
rescale each longitudinal value, multiply the respective week's rescaling ratio
by the cross-sectional value. Rescaled longitudinal data was compared against
time-series data for multiple test counties and was equivalent. For a more
in-depth explanation of this procedure, see \citet[p. 5]{park_etal}. Missing
datapoints in longitudinal datasets were filled in with interpolated values
using \texttt{zoo\:\:na.approx()} \citep{zoo}.

@@

\subsection{Analysis}
Google Search Trends data and additional demographic data were merged with each
individual indicator based on County FIPS Codes and date using \texttt{dplyr} in \texttt{r}, version 4.1.2 \citep{tidyverse}. After creating these
different datasets, I use the Pearson correlation formula (formula \eqref{eq:pearsoncorr}) for cross-sectional numeric data to calculate the
strength of the relationship between each Google Trend and the respective
data source.

@@

To address longitudinal correlations, I employ Repeated Measures Correlation
using the `rmcorr` package in \texttt{r} \citep{bland1995, bakdash2017}. 
Repeated Measures Correlation is useful for determining the within-county
association for paired measures across time and counties. Results of these 
correlations can be seen in table \ref{tab:corr-results}.

As an additional test of the relationships, I employ multiple linear regression
(for cross-sectional datasets) and random intercept hierarchical linear models
\citep{pinheiro_etal21} (for longitudinal datasets) to identify the strength of
relationships across locales. I include city-data like county population size,
broadband rates, and median income to attempt to disentangle possible
confounders of the relationship between Google search Trends and verified
indicators. Identifying these possible confounded relationships will help to
explain why some articles find relationships between the trends and outcomes
while others did not. For linear regression models, I normalize independent
variables, i.e.variables have been centered and scaled to have a mean of 0 and
standard deviation 1. Results of these models can be seen in tables \ref{tab:vacc_hes_analysis} through \ref{tab:pres_2020_analysis}.

### 2022-03-04 paper2.tex
finish paper 1

\begin{figure}
{\centering \includegraphics[width=0.8\linewidth]{figs/paper2/plot-model-1-1}}
\caption{Plot of Coefficients, Model 1}\label{fig:plot-model-1}
\end{figure}

@@

\begin{figure}
{\centering \includegraphics[width=0.8\linewidth]{figs/paper2/plot-model-3-1}}
\caption{Plot of Coefficients, Model 3}\label{fig:plot-model-3}
\end{figure}

>>>>

\begin{figure}[h]
{\centering \includegraphics[width=0.8\linewidth]{figs/paper2/plot-model-1-1}}
\caption{What predicts someone intentionally searched for information, Plot of Coefficients}\label{fig:plot-model-1}
\end{figure}

@@

% <!-- TODO: watch out about saying vaccinated, this variable is about opinion -->

\begin{figure}[h]
{\centering \includegraphics[width=0.8\linewidth]{figs/paper2/plot-model-3-1}}
\caption{Does receiving or searching for information predict vaccination status? Plot of Coefficients}\label{fig:plot-model-3}
\end{figure}

### 2022-03-04 paper3.tex
finish paper 1

@@

\begin{figure}
{\centering \includegraphics[width=0.8\linewidth]{figs/paper3/model}}
\caption{Preliminary model of norm formation}\label{fig:norm-framework}
\end{figure}

@@

To estimate the concept of `real threat of infection' in figure
\ref{fig:norm-framework}, the models use county-level Covid-19 case rates with
data compiled by The New York Times \citeyearpar{covid_data}. It is widely acknowledged that there are biases in this data due to inconsistencies and availability in testing as well as
different community propensity to test \citep{gu22, cdc20a}. However, it is the best
measure we have of actual case rates. County data are scaled with a rolling mean
with a rolling window width of seven days to smooth out intra-week noise. Case
Rate is measured as number of cases per 100,000 population. Observations vary
from a minimum of 0 to a maximum of 1,565 across the two datasets.

>>>>

% need to explain a little bit more what diversity of signals etc means (since this is the first look)

@@

@@

To estimate the concept of `real threat of infection' for a fear-based model of
health behaviors, the models use county-level Covid-19 case rates with
data compiled by The New York Times \citeyearpar{covid_data}. It is widely acknowledged that there are biases in this data due to inconsistencies and availability in testing as well as
different community propensity to test \citep{gu22, cdc20a}. However, it is the best
measure we have of actual case rates. County data are scaled with a rolling mean
with a rolling window width of seven days to smooth out intra-week noise. Case
Rate is measured as number of cases per 100,000 population. Observations vary
from a minimum of 0 to a maximum of 1,565 across the two datasets.
